---
title: "A3"
author: "Braden Baker"
date: "3/31/2019"
output: github_document
---

## Problem 1
I have placed a copy of the Ames Iowa real estate data in Moodle. Download this file and load it into your R session. Then run str() on it to make sure you have the data and know the names of the variables.

```{r}
#Place your R code here.
library(tidyverse)
library(openintro)
load("A3workspace.RData")

str(ames)

```

## Problem 2
Use the code in the document "Train Test" in Moodle to create Train and Test subsets based on random sampling with a 75/25 split. Verify that the numbers of observations in Train and Test add up to the number of observations in ames and that the split is correct.
```{r}
#Place your R code here.
nrows = nrow(ames)

set.seed(999)

rows <- sample(nrows)

ames = ames[rows,]

head(ames)

split <- round(nrow(ames) * .75)

trainA = ames[1:split,]
testA = ames[(split+1):nrows,]

nrow(trainA)
nrow(testA)
nrow(ames)

```

## Problem 3
Select any three variables to serve as predictors of sale price. Build a regression model based on Train using these variables. Examine the coefficient signs and p-values to make sure you have a model that would earn the approval of an econometrician. If the results are not satisfactory, try some other variables.

Finally, display the summary, and report your residual standard error and R-squared.

```{r}
#Place your R code here.
#Bedroom.AbvGr

modA <- lm(SalePrice~Fireplaces+ Lot.Area + Year.Built, data=trainA)
summary(modA)
```


## Problem 4
Use predict() to build a vector pred for the cases in Test. Create measures of performance to match the residual standard error and R-Square from the Train results. Use these four numbers to compare how well the model performed on Test with how well it performed on Train.

```{r}
#Place your R code here.

predA <- predict(modA, newdata = testA)

Review <- cbind(testA,predA) %>% 
  select(SalePrice, predA, Lot.Area, Year.Built, Fireplaces ) %>% 
  mutate(res = SalePrice - predA)
head(Review)

sstot = sum((Review$SalePrice - mean(Review$SalePrice))^2)
ssres = sum(Review$res^2)
rsq = 1 - (ssres/sstot)
rsq
ResStandErrorPredA <- sd(Review$res)
ResStandErrorPredA

summary(Review$predA)
summary(Review$SalePrice)
```