---
title: "A5"
author: "Braden Baker"
date: "April 16, 2019"
output: github_document
---

## Problem 1
Load the cdc dataset.  Review my document "Data Analysis Example" from Canvas.  Use the cleaning and recoding I did to produce a clean and enhanced dataset for analysis. Do a summary to verify the correctness of your work.

```{r}
# Place your code here. Include load and library commands.
library(tidyverse)
library(gmodels)
library(caret)

load("cdc.Rdata")

cdc$BMI = (cdc$weight*703)/(cdc$height)^2
cdc$BMIDes = (cdc$wtdesire*703)/(cdc$height)^2
cdc$DesActRatio = cdc$BMIDes/cdc$BMI
cdc$BMICat = cut(cdc$BMI,c(18,5,24.9,29.9,39.9,200),labels = 
       c("Underweight","Normal","Overweight",
       "Obese","Morbidly Obese"),include.lowest=T)
cdc$BMIDesCat = cut(cdc$BMIDes,c(18,5,24.9,29.9,39.9,200),labels = 
       c("Underweight","Normal","Overweight",
       "Obese","Morbidly Obese"),include.lowest=T)
cdc$ageCat = cut_number(cdc$age,n=4,labels=c("18-31","32-43","44-57","58-99"))

cdcmale <- cdc %>%
  filter(gender=="m") %>%
  mutate(gender=1)

  

cdcfemale <- cdc %>%
  filter(gender=="f") %>%
  mutate(gender=0)
cdc <- rbind(cdcmale, cdcfemale)
summary(cdc)
head(cdc)
tail(cdc)
```


## Problem 2
Split the dataframe into train and test with a 70/30 ratio.  Do an str() on both train and test to verify that you have what you want.

```{r}
# Place your code here.

nrows = nrow(cdc)

set.seed(999)

rows <- sample(nrows)

cdc = cdc[rows,]

split <- round(nrow(cdc) * .70)

traincdc = cdc[1:split,]
testcdc = cdc[(split+1):nrows,]

nrow(traincdc)
nrow(testcdc)
nrow(cdc)
str(traincdc)
str(testcdc)
```


## Problem 3
Create a logistic regression model which predicts gender based on height and weight using train. Measure its accuracy with a confusion matrix based on test.

```{r}
# Place your code here.
cdcglm <- glm(gender~height+weight, data = traincdc, family = "binomial")
summary(cdcglm)

predcdc <- predict(cdcglm, newdata = testcdc, type = "response")
predcdc <- predcdc > .5


matrix <- table(testcdc$gender,predcdc)
matrix
accuracy <- sum(matrix[1,1]+matrix[2,2])/sum(matrix[1,1]+matrix[1,2]+matrix[2,1]+matrix[2,2])
accuracy
```


## Problem 4
Follow the same process as in Problem 3, but add the exercise and smoking variables. Did this improve on the accuracy you achieved in Problem 3.


This did improve the score.

```{r}
# Place your code here.
cdcglm <- glm(gender~height+weight + exerany + smoke100, data = traincdc, family = "binomial")
summary(cdcglm)

predcdc <- predict(cdcglm, newdata = testcdc, type = "response")
predcdc <- predcdc > .5


matrix <- table(testcdc$gender,predcdc)
matrix
accuracy <- sum(matrix[1,1]+matrix[2,2])/sum(matrix[1,1]+matrix[1,2]+matrix[2,1]+matrix[2,2])
accuracy
```


## Problem 5
One of the conclusions we reached in the original analysis is that women were more anxious to keep their weight low than men.  Formulate a variable that captures this and add it to the model. Explain the formulation you chose. Add this variable to the model in Problem 4. Did it improve your accuracy?


I created the variable, wtdesireddif, which is the difference between weight desired and actual weight. I chose this as a variable because women tend to want to keep their weight lower, wtdesireddif would be a good way of capturing that as women will tend to have negative values more often than men. After adding this variable to the model the accuracy did improve. 

```{r}
# Place your code here.

traincdc$wtdesiredif <- traincdc$wtdesire - traincdc$weight
testcdc$wtdesiredif <- testcdc$wtdesire - testcdc$weight

cdcglm <- glm(gender~height+weight + exerany + smoke100 + wtdesiredif, data = traincdc, family = "binomial")
summary(cdcglm)

predcdc <- predict(cdcglm, newdata = testcdc, type = "response")
predcdc <- predcdc > .5


matrix <- table(testcdc$gender,predcdc)
matrix
accuracy <- sum(matrix[1,1]+matrix[2,2])/sum(matrix[1,1]+matrix[1,2]+matrix[2,1]+matrix[2,2])
accuracy
```
