---
title: "A6"
author: "Braden Baker"
date: "4/26/2019"
output: github_document
---

This assignment will make use of a published tutorial to build a knn model of gender classification for the cdc data using the caret package.

## Load required libraries and the data.

```{r}
library(caret)
library(e1071)
```

## Problem 1

Load the cdc data, do the cleaning and transformations. Do not do the splitting. Run str() and summary to verify your results.

```{r}
load("cdc.Rdata")

cdc$BMI = (cdc$weight*703)/(cdc$height)^2
cdc$BMIDes = (cdc$wtdesire*703)/(cdc$height)^2
cdc$DesActRatio = cdc$BMIDes/cdc$BMI
cdc$BMICat = cut(cdc$BMI,c(18,5,24.9,29.9,39.9,200),labels = 
       c("Underweight","Normal","Overweight",
       "Obese","Morbidly Obese"),include.lowest=T)
cdc$BMIDesCat = cut(cdc$BMIDes,c(18,5,24.9,29.9,39.9,200),labels = 
       c("Underweight","Normal","Overweight",
       "Obese","Morbidly Obese"),include.lowest=T)
cdc$ageCat = cut_number(cdc$age,n=4,labels=c("18-31","32-43","44-57","58-99"))

str(cdc)
```

## Problem 2
Use caret to split the data into training and testing dataframes with a 70/30 split. Use set.seed(2343) to facilitate replicability. Run dim() on both to verify your results.

```{r}
nrows = nrow(cdc)

rows <- sample(nrows)

cdc = cdc[rows,]


split <- round(nrow(cdc) * .7)

traincdc = cdc[1:split,]
testcdc = cdc[(split+1):nrows,]

dim(traincdc)
dim(testcdc)
```


## Problem 3

Build a knn model, knn_fit, using height and weight to predict gender using the train function in caret. in the call to trainControl reduce number to 5 and repeats to 2. Then use the remainder of the code in the example to compute the confusion matrix and accuracy for this model.

```{r}
trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 2)

knn_fit <- train(gender ~ height+weight, data = traincdc, method = "knn",
 trControl=trctrl,
 preProcess = c("center", "scale"),
 tuneLength = 10)

predcdc <- predict(knn_fit, newdata = testcdc)
matrix <- table(testcdc$gender,predcdc)
matrix
accuracy <- sum(matrix[1,1]+matrix[2,2])/sum(matrix[1,1]+matrix[1,2]+matrix[2,1]+matrix[2,2])
accuracy
```

## Problem 4
Repeat the process in Problem 3, adding exerany and smoke100 to the model.  Did this improve the accuracy of the model?

This did improve the accuracy slighty.

```{r}
# Place your code here.
trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 2)

knn_fit <- train(gender ~ height+weight+exerany+smoke100, data = traincdc, method = "knn",
 trControl=trctrl,
 preProcess = c("center", "scale"),
 tuneLength = 10)

predcdc <- predict(knn_fit, newdata = testcdc)
matrix <- table(testcdc$gender,predcdc)
matrix
accuracy <- sum(matrix[1,1]+matrix[2,2])/sum(matrix[1,1]+matrix[1,2]+matrix[2,1]+matrix[2,2])
accuracy
```

## Problem 5.
Repeat Problem 4, adding a variable of your choice which captures the extra sensitivity of women to weight. Does this improve the accuracy?

This improved the accuracy of the model, more than adding smoke100 and exerany.

```{r}
# Place your code here.
traincdc$wtdesiredif <- traincdc$wtdesire - traincdc$weight
testcdc$wtdesiredif <- testcdc$wtdesire - testcdc$weight

trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 2)

knn_fit <- train(gender ~ height+weight+exerany+smoke100+wtdesiredif, data = traincdc, method = "knn",
 trControl=trctrl,
 preProcess = c("center", "scale"),
 tuneLength = 10)

predcdc <- predict(knn_fit, newdata = testcdc)
matrix <- table(testcdc$gender,predcdc)
matrix
accuracy <- sum(matrix[1,1]+matrix[2,2])/sum(matrix[1,1]+matrix[1,2]+matrix[2,1]+matrix[2,2])
accuracy
```

